---
title: "Homework2"
output: github_document
---

# Homework 2

Here is an exploration of the *Gapminder* dataset using `ggplot2` and `dplyr`.

### 1. First, the packages must be loaded.

*Gapminder* is the dataset itself, while `tidyverse` is the meta-package, a collection with tools like `ggplot2` for graphics, `tbl_df` for tibble data.frames, and `dplyr` for manipulating data.

```{r}

library(gapminder)
library(tidyverse)

```

### 2. What kind of dataset is *Gapminder*?

A "data.frame" of demographic metrics by countries and how they relate to each other over time. The summary function shows that there are 6 columns. Two are composed of frequency counts (`country` and `continent`), and four are numerical quantities (`year`, `lifeExp`, `pop`, and `gdpPercap`). These values span 1704 rows.

```{r}

summary(gapminder)
class(gapminder)
ncol(gapminder)
nrow(gapminder)

```

What years does the *Gapminder* dataset consider? It spans 55 years, from 1952 to 2007.

```{r}

max(gapminder$year)-min(gapminder$year)

```

Unlike a matrix that is composed only of numerical values, a "data.frame" can have more than one type of data, like integers and doubles. *Gapminder* has both, so it is a "data.frame".

```{r}

sapply(gapminder, typeof)

```

### 3. Exploring the life expectancy of Chile over time.

If you plot Chile's life expactancy `lifeExp` over time `year`, and make the points look like mountains, you get the Andes cordillera range!

```{r}

filter(gapminder, country == "Chile") %>%
ggplot(aes(year, lifeExp)) +
  geom_point(shape = 2, stroke = 2, size = 4)

```

We can also compare all the countries that share the Andes mountain range with Chile.

```{r}
filter(gapminder, country == "Chile"
       | country == "Argentina"
       | country == "Peru"
       | country == "Bolivia") %>% 
ggplot(aes(year,lifeExp)) +
  geom_point(aes(colour = country), shape = 2, stroke = 2, size = 4)
```

We can also look at `lifeExp` by itself through "violin-and-jitter" plots.

```{r}

filter(gapminder, country == "Chile"
       | country == "Argentina"
       | country == "Peru"
       | country == "Bolivia") %>% 
  ggplot(aes(country, lifeExp)) +
  geom_violin() +
  geom_jitter(alpha=0.2, shape=1, size=5, stroke=2)

```

In contrast, we can also look at `pop` or population through "violin-and-jitter" plots.

```{r}

filter(gapminder, country == "Chile"
       | country == "Argentina"
       | country == "Peru"
       | country == "Bolivia") %>% 
  ggplot(aes(country, pop)) +
  geom_violin() +
  geom_jitter(alpha=0.2, shape=1, size=5, stroke=2)

```

### 4. Exploring the raw data of Chile's neighbours.

```{r}

filter(gapminder, country == c("Chile", "Argentina", "Bolivia", "Peru"))

```

I was born in Chile in 1993, so I'm interested in summary data of the region during that time. *Gapminder* has the year 1992, which is close enough.

```{r}

filter(gapminder,
       
       country == "Chile"
       | country == "Argentina"
       | country == "Peru"
       | country == "Bolivia",
       
       year == 1992) %>%
  arrange(desc(gdpPercap)) %>% 
  select(country, lifeExp, pop, gdpPercap) %>% 
  knitr::kable(digits = 0)

```

I guess I can expect to live to 74 years old!
